---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# yatah

<!-- badges: start -->
```{r, echo = FALSE}
version <- as.vector(read.dcf('DESCRIPTION')[, 'Version'])
version <- gsub('-', '.', version)
```
[![license](https://img.shields.io/badge/license-GPL--3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0.en.html)
[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![packageversion](https://img.shields.io/badge/Package%20version-`r version`-orange.svg?style=flat-square)](commits/master)
[![Travis build status](https://travis-ci.org/abichat/yatah.svg?branch=master)](https://travis-ci.org/abichat/yatah)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/abichat/yatah?branch=master&svg=true)](https://ci.appveyor.com/project/abichat/yatah)
[![Codecov test coverage](https://codecov.io/gh/abichat/yatah/branch/master/graph/badge.svg)](https://codecov.io/gh/abichat/yatah?branch=master)
[![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-yellowgreen.svg)](/commits/master)
<!-- [![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/yatah)](https://cran.r-project.org/package=yatah) -->
<!-- badges: end -->

The goal of __yatah__ is to manage taxonomy when lineages are described with strings and ranks separated with special patterns like `|*__` or `;*__`.

For instance, the well-known _Escherichia coli_ could be coded as `k__Bacteria|p__Proteobacteria|c__Gammaproteobacteria|o__Enterobacteriales|f__Enterobacteriaceae|g__Escherichia|s__Escherichia_coli`.

## Installation

<!-- You can install the released version of yatah from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("yatah") -->
<!-- ``` -->

<!-- And the development version from [GitHub](https://github.com/) with: -->
You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("abichat/yatah")
```
## Example

```{r}
library(yatah)
```


```{r}
lineages <- c(
  "k__Bacteria|p__Actinobacteria|c__Actinobacteria|o__Coriobacteriales",
  "k__Bacteria|p__Bacteroidetes|c__Bacteroidia|o__Bacteroidales",
  "k__Bacteria|p__Bacteroidetes|c__Flavobacteriia|o__Flavobacteriales",
  "k__Bacteria|p__Firmicutes|c__Bacilli|o__Bacillales",
  "k__Bacteria|p__Firmicutes|c__Bacilli|o__Lactobacillales",
  "k__Bacteria|p__Firmicutes|c__Clostridia|o__Clostridiales",
  "k__Bacteria|p__Proteobacteria|c__Epsilonproteobacteria|o__Campylobacterales",
  "k__Bacteria|p__Proteobacteria|c__Gammaproteobacteria|o__Enterobacteriales",
  "k__Bacteria|p__Proteobacteria|c__Gammaproteobacteria|o__Pseudomonadales"
)
```

* `is_rank()` checks if the lineages are of the desired rank.

```{r is_rank}
is_rank(lineages, rank = "order")
is_rank(lineages, rank = "species")
```

* `is_clade()` checks if the lineages belong to the desired clade.

```{r is_clade}
is_clade(lineages, clade = "Proteobacteria", rank = "phylum")
```


* `last_clade()` extracts the last clade of the lineages.

```{r last_rank}
last_clade(lineages)
```

* `taxtable()` computes the taxonomic table corresponding to the lineages.

```{r taxtable}
table <- taxtable(lineages)
table
```

* `taxtree()` computes the taxonomic tree (format `phylo`) from a taxonomic table. 

```{r taxtree}
tree <- taxtree(table)
tree
plot(tree, show.node.label = TRUE)
```

