---
title: "Taxonomy of Gammaproteobacteria"
output: rmarkdown::html_vignette
bibliography: references.bib
link-citations: yes
vignette: >
  %\VignetteIndexEntry{Taxonomy_of_Gammaproteobacteria}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Preliminaries

This vignette illustrates the most useful functions of **yatah**. 

```{r setup, message=FALSE}
library(dplyr)
library(yatah)
```

## Data

For this example, we use data from @zeller2014potential. It is the abundances of bacteria present in 199 stool samples.

```{r}
abundances <- as_tibble(yatah::abundances)
print(abundances, n_extra = 2)
```

```{r data, message=FALSE}
taxonomy <- select(abundances, lineages)
taxonomy
```

## Filtering

Here, we have all the present bacteria at all different ranks. As we are just interested in genera that belong to the _Gammaproteobacteria_ class, we `filter()` the lineages with `is_clade()` and `is_rank()`. The genus name is accessible with `last_clade()`.

```{r filter}
gammap_genus <-
  taxonomy %>% 
  filter(is_clade(lineages, "Gammaproteobacteria"),
         is_rank(lineages, "genus")) %>% 
  mutate(genus = last_clade(lineages))
gammap_genus
```

## Taxonomic table

It is useful to have a taxonomic table. `taxtable()` do the job.

```{r table}
gammaprot_table <-
  gammap_genus %>% 
  pull(lineages) %>% 
  taxtable()
as_tibble(gammaprot_table)
```

## Taxonomic tree

To have a tree, use `taxtree()` with a taxonomic table in input. By default, it collapses ranks with only one subrank.

```{r tree}
gammaprot_tree <- taxtree(gammaprot_table)
gammaprot_tree
```


```{r ggtree, fig.width=7, fig.height=7}
plot(gammaprot_tree, show.node.label = TRUE, cex = 0.7, 
     main = "Taxonomy of Gammaproteobacteria")
```

## References
